---
title: 'Reproducible Research: Peer Assessment 1'
author: "Schmaef"
date: '2022-07-14'
output:
  html_document:
    keep_md: yes
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data

The Data is downloaded from my github repo and read into the variable "activity.raw". Also some packages are loaded.
```{r loading data}
library(lattice)
library(ggplot2)

download.file("https://github.com/Schmaef/RepData_PeerAssessment1/blob/master/activity.zip", destfile="activity.zip", method="curl")
unzip ("activity.zip")

activity.raw <- read.csv("activity.csv", sep =",")
```

## What is mean total number of steps taken per day?

To calculate the mean total number of steps, steps and data are aggregated summed up
```{r date}
daily.steps <- aggregate(steps ~ date, activity.raw, sum)
```

A histogramm is created and a few optical variables are changed
```{r plot}
hist(daily.steps$steps,
     main = "total daily steps",
     xlab = "total steps",
     breaks=20)
```

Lastly the mean, median and total are calculated
```{r mean median total}
mean(daily.steps$steps)

median(daily.steps$steps)

sum(daily.steps$steps)
```

## What is the average daily activity pattern?

To calculate the average daily activity pattern, the steps and interval are aggregated and averaged
```{r int.steps}
int.steps <- aggregate(steps ~ interval, activity.raw, mean)
```

The data is presented in a time series plot
```{r plot2}
xyplot(steps ~ interval, int.steps, 
       type = "l",
       main = "average daily activity pattern",
       xlab = "interval", 
       ylab = "steps")
```

The interval with the maximum average steps is calculated
```{r interval}
int.steps$interval[which.max(int.steps$steps)]
```

## Imputing missing values

To impute missing data, blanks are filled with the average steps per interval
```{r imputing}
activity.nn <- activity.raw
activity.nn[is.na(activity.nn$steps), "steps"] = mean(int.steps$steps) 
```

Repeating the code from the first histogram
```{r date 2}
daily.steps.nn <- aggregate(steps ~ date, activity.nn , sum)

hist(daily.steps.nn$steps,
     main = "total daily steps with imputed data",
     xlab = "total steps",
     breaks=20)
```

Again the mean, median and total are calculated
```{r mean median}
mean(daily.steps.nn$steps)

median(daily.steps.nn$steps)

sum(daily.steps.nn$steps)
```

## Are there differences in activity patterns between weekdays and weekends?

To seperate between weekdays and weekend a new factor variable called daytype is created

```{r weekdays}
activity.nn$day <- as.POSIXlt(activity.nn$date)$wday
activity.nn$daytype <- as.factor(ifelse(
  activity.nn$day == 0 | activity.nn$day == 6, "weekend", "weekday"))
activity.nn <- subset(activity.nn, select = -c(day))
```

The data is presented in two different plots, one for weekdays and one for weekends,
as the daytype variable defines. "layout = c(1,2)" is used to get the two plots into one window 

```{r weekdays plot}
xyplot(steps ~ interval | activity.nn$daytype, int.steps, 
       type = "l", 
       layout = c(1,2),
       main = "average daily activity pattern",
       xlab = "interval", 
       ylab = "steps")
```
